Parity games in pushdown systems with parameters

### 2023-07-04 reachability/article.tex
Too long, 13 pages and something, have to shorten to 12

Our main result consists in
proving parametric pushdown reachability games
and parametric pushdown parity games belong to $(n+1)$-$\EXP$ in case the number of parameters
$n$ is fixed,
and providing a nonelementary lower bound for parametric pushdown reachability games in general.

@@

	For the nonelementary lower bound,
we reduce the $\FO$ satisfiability problem on words, known to be nonelementary
from \cite{Sto74}, to the
problem of deciding whether a player has a winning strategy for parametric pushdown reachability games.

@@

We first consider {\em reachability games}: 
the winning condition $\win_F(S)$ is given by a set of final configurations
$F \subseteq S$, and $\win_F(S)$ is the set of maximal plays in $S^* F S^\infty$. 
Secondly, we consider {\em parity games}: 
the winning condition $\win_\Omega(S)$ is given by a {\em priority function} 
 $\Omega : S \to  [0, m]$.
An infinite path $ \pi = s_0 s_1 s_2 \ldots$ is in $\win_\Omega(S)$ if the smallest priority appearing infinitely often in
the sequence 
 is even, i.e.
if $\min( \{ \Omega(s) | ~ s \in \text{\textit{Inf}}( \pi) \})$ is even.  
A finite play is in $\win_\Omega(S)$ if its target configuration is in $S_1$.
We simply write
$\win_F$  and $\win_\Omega$ in case the set $S = S_0 \biguplus S_1$ is obvious from context.

@@

In the particular case where $P=\{p\}$ is a singleton for some parameter $p$ and $\mu(p)=u \in \Gamma^*$,
we prefer to write $q(w)\rightarrow_{\mathcal{Z}, u}q'(w')$
to denote $q(w)\rightarrow_{\mathcal{Z}, \mu}q'(w')$. In case the automaton $\mathcal{Z}$ is obvious from context, we write $\rightarrow_{\mu}$ (resp. $\rightarrow_{u}$)
instead of $\rightarrow_{\mathcal{Z}, \mu}$ (resp. $\rightarrow_{\mathcal{Z},u}$). 
Since we are concerned with the graph of configurations and not the language recognized, we omit the input alphabet from the model and will use the name pushdown system rather than pushdown automaton.

\subsection{PPDS games}

We are interested in this article in parametric variants of reachability games and parity games.
One goal may be to characterize the set of all possible parameter valuations for which a player has a winning strategy in a 
reachability game or a parity game
played on the arena induced by the
resulting concrete automaton
and some partition of its set of states. 
For determining such a set of parameter valuations, we incorporate the choice of the parameter valuation to the game. Each player then wants the parameter valuation to lead to an arena where it has a winning strategy.

@@

\problemx{Parametric pushdown reachability game}
{A parametric pushdown automaton $\mathcal{Z}= (Q, \Gamma, \{p_0, p_1, \ldots, p_k \}, R, q_{init},  \gamma_{init} , F)$, where $Q = Q_0  \biguplus Q_1 $, and $ P = P_0 \biguplus P_1$.}
{Does player $0$ have a winning strategy from $\mu_\bot$ in the reachability game \\
$\mathcal{G} =(A_{(\mathcal{Z}, P_0, P_1, Q_0 , Q_1)}, \win_{F \times \Gamma^* \times (\Gamma^*)^P})$ ?\newline}

@@

We write {\sc $n$-parametric reachability parity game}
and
{\sc $n$-parametric pushdown parity game} if 
 the number of parameters $n=|P|$ of $\mathcal{Z}$ is fixed by the problem.
Remark that the parametric reachability problem for an automaton with parameters can be seen as solving the parametric reachability game where $P_1 = \emptyset$, and $Q_1 = \emptyset$.

@@

Note
{\sc $n$-parametric pushdown parity game} 
is polynomial time reducible to
{\sc $n$-pebble pushdown parity game}.
See appendix for more details.

@@

For solving parametric pushdown parity games, on the other hand, we introduced the notion of pebble pushdown parity games, and
pebble pushdown automata over unary alphabet plus a bottom-of-stack symbol
can be seen as a form of one-counter automata extended with pebbles. 
A pebble one-counter automata is then a one-counter automata that
can use a set of pebbles as markings on the set of non-negative integers, with a 
$\text{\textit{drop}}$ recording the current counter value, and a 
$\text{\textit{lift}}$ poping the last dropped
pebble while requiring that the current counter value 
corresponds to the one from the last dropped pebble.
%
Unlike for parametric extensions, 
decidability of 
what should be called
pebble one-counter parity games follows from
decidability of pebble pushdown parity games. 
We believe it natural to study such a pebble one-counter automata model
on its own.
In particular we believe that the question of whether or not
solving pebble one-counter reachability games is nonelementary is worth investigating.

Of note is that reachability games and parity games are 
not the only problems one can consider, as one can
also explore the complexity of reachability itself. 
In particular
one can 
study the precise complexity of the
{\sc $n$-PPDA reachability} problem.
Since reachability can be viewed as a reachability game, we know that
{\sc $n$-PPDA reachability} is decidable in $(n+1)$-$\EXP$, 
but the precise complexity of the problem remains unknown.
We are furthermore convinced that it is worthwhile to investigate further comparisons between parametric pushdown automata, pebble pushdown automata, 
and register pushdown automata, especially since the latter have been shown to only really ``remember'' at most  $3r$ data values, where $r$ is the number of registers.

>>>>

Our main result consists in
proving parametric pushdown parity games belong to $(n+1)$-$\EXP$ in case the number of parameters
$n$ is fixed,
and providing a nonelementary lower bound for parametric pushdown parity games in general.

@@

	For the nonelementary lower bound,
we reduce the $\FO$ satisfiability problem on words, known to be nonelementary
from \cite{Sto74}, to the
problem of deciding whether a player has a winning strategy for parametric pushdown parity games.

@@

We consider {\em parity games}: 
the winning condition $\win_\Omega(S)$ is given by a {\em priority function} 
 $\Omega : S \to  [0, m]$.
An infinite path $ \pi = s_0 s_1 s_2 \ldots$ is in $\win_\Omega(S)$ if the smallest priority appearing infinitely often in
the sequence 
 is even, i.e.
if $\min( \{ \Omega(s) | ~ s \in \text{\textit{Inf}}( \pi) \})$ is even.  
A finite play is in $\win_\Omega(S)$ if its target configuration is in $S_1$.
We simply write
$\win_\Omega$ in case the set $S = S_0 \biguplus S_1$ is obvious from context.

@@

\textbf{PPDS games.}
We are interested in this article in parametric variants of parity games.
One goal may be to characterize the set of all possible parameter valuations for which a player has a winning strategy in a parity game
played on the arena induced by the
resulting concrete automaton
and some partition of its set of states. 
For determining such a set of parameter valuations, we incorporate the choice of the parameter valuation to the game. Each player then wants the parameter valuation to lead to an arena where it has a winning strategy.

@@

@@

We write 
{\sc $n$-parametric pushdown parity game} if 
 the number of parameters $n=|P|$ of $\mathcal{Z}$ is fixed by the problem.

@@

@@

For solving parametric pushdown parity games, on the other hand, we introduced the notion of pebble pushdown parity games, and
pebble pushdown automata over unary alphabet plus a bottom-of-stack symbol
can be seen as a form of one-counter automata extended with pebbles. 
A pebble one-counter automata is then a one-counter automata that
can use a set of pebbles as markings on the set of non-negative integers, with a 
$\text{\textit{drop}}$ recording the current counter value, and a 
$\text{\textit{lift}}$ poping the last dropped
pebble while requiring that the current counter value 
corresponds to the one from the last dropped pebble.
%
Unlike for parametric extensions, 
decidability of 
what should be called
pebble one-counter parity games follows from
decidability of pebble pushdown parity games. 
We believe it natural to study such a pebble one-counter automata model
on its own.
In particular we believe that the question of whether or not
solving pebble one-counter parity games is nonelementary is worth investigating.

Of note is that parity games are 
not the only problems one can consider, as one can
also explore the complexity of reachability. 
In particular
one can 
study the precise complexity of the
{\sc $n$-PPDA reachability} problem.
Since reachability can be viewed as a reachability game, which can be simulated using a parity game, we know that
{\sc $n$-PPDA reachability} is decidable in $(n+1)$-$\EXP$, 
but the precise complexity of the problem remains unknown.
We are furthermore convinced that it is worthwhile to investigate further comparisons between parametric pushdown automata, pebble pushdown automata, 
and register pushdown automata, especially since the latter have been shown to only really ``remember'' at most  $3r$ data values, where $r$ is the number of registers.
